Assessing Read Quality and Trimming

### 2023-05-21 episodes/quality-control.Rmd
Initial commit seq assembly episode.
Fixed the snippy part to avoid massive downloads.

```output
ERR029206_1.fastq.gz                   ERR029207_2.fastq.gz
ERR029206_2.fastq.gz                   GCF_000195955.2_ASM19595v2_genomic.fna
ERR029207_1.fastq.gz
```

@@

@@

Yesterday we downloaded 2 samples. How many FASTQ files are in this dataset? Why is this? What is the difference between the two files that start with the same name? Discuss with your neighbour.

@@

There are 4 FASTQ files. Each sample has [paired-end reads](https://thesequencingcenter.com/knowledge-base/what-are-paired-end-reads/).

@@

In total, it should take about five minutes for FastQC to run on all
four of our FASTQ files. When the analysis completes, your prompt
will return. So your screen will look something like this:

```output
Approx 85% complete for ERR029207_2.fastq.gz
Approx 90% complete for ERR029207_2.fastq.gz
Approx 95% complete for ERR029207_2.fastq.gz
Analysis complete for ERR029207_2.fastq.gz
data $
```

@@

```output
ERR029206_1.copy.fastq                 ERR029207_1.fastq.gz
ERR029206_1.fastq.gz                   ERR029207_1_fastqc.html
ERR029206_1_fastqc.html                ERR029207_1_fastqc.zip
ERR029206_1_fastqc.zip                 ERR029207_2.fastq.gz
ERR029206_2.fastq.gz                   ERR029207_2_fastqc.html
ERR029206_2_fastqc.html                ERR029207_2_fastqc.zip
ERR029206_2_fastqc.zip                 GCF_000195955.2_ASM19595v2_genomic.fna
```

@@

```bash
$ firefox ERR026473_1_fastqc.html
```

@@

Discuss your results with a neighbor. Which sample(s) looks the best
in terms of per base sequence quality? Which sample(s) look the
worst?

All samples are of reasonable quality.

@@

```output
ERR029206_1.copy.fastq                 ERR029207_1.fastq.gz
ERR029206_1.fastq.gz                   ERR029207_1.fastq.gz_trim.fastq
ERR029206_1.fastq.gz_trim.fastq        ERR029207_2.fastq.gz
ERR029206_1_trim.fastq                 ERR029207_2.fastq.gz_trim.fastq
ERR029206_2.fastq.gz                   GCF_000195955.2_ASM19595v2_genomic.fna
ERR029206_2.fastq.gz_trim.fastq
```

@@

```output
ERR029206_1.fastq.gz_trim.fastq ERR029207_1.fastq.gz_trim.fastq
ERR029206_2.fastq.gz_trim.fastq ERR029207_2.fastq.gz_trim.fastq
```

>>>>

```output
ERR029206_1.fastq.gz
ERR029206_2.fastq.gz 
```

@@

You've already seen that part, but repetition doesn't hurt. If you feel that it's anyway unnecessary, move on to the next part.

@@

We downloaded the samples. How many FASTQ files are in this dataset? Why is this? What is the difference between the two files that start with the same name? Discuss with your neighbour.

@@

There are 2 FASTQ files. Each sample has [paired-end reads](https://thesequencingcenter.com/knowledge-base/what-are-paired-end-reads/).

@@

In total, it should take about 15-20 minutes for FastQC to run on all
four of our FASTQ files. When the analysis completes, your prompt
will return. So your screen will look something like this:

```output
Approx 85% complete for ERR029206_2.fastq.gz
Approx 90% complete for ERR029206_2.fastq.gz
Approx 95% complete for ERR029206_2.fastq.gz
Analysis complete for ERR029206_2.fastq.gz
data $
```

@@

```output
...
ERR029206_1.copy.fastq
ERR029206_1.fastq.gz
ERR029206_1_fastqc.html
ERR029206_1_fastqc.zip
ERR029206_2.fastq.gz
ERR029206_2_fastqc.html
ERR029206_2_fastqc.zip
```

@@

```bash
$ firefox ERR029206_1_fastqc.html
```

@@

Discuss your results with a neighbor. Do the sample looks good in terms of per base sequence quality?

@@

```output
...
ERR029206_1.copy.fastq 
ERR029206_1.fastq.gz
ERR029206_1.fastq.gz_trim.fastq 
ERR029206_1_trim.fastq 
ERR029206_2.fastq.gz
ERR029206_2.fastq.gz_trim.fastq
GCF_000195955.2_ASM19595v2_genomic.fna
```

@@

```output
...
ERR029206_1.fastq.gz_trim.fastq 
ERR029206_2.fastq.gz_trim.fastq 
```
